---
title: "Assignment 2"
subtitle: Group 85
date: "20/03/2026"
output:
  pdf_document: default
  html_document:
    df_print: paged
author:
  - "Dovydas Venclovas - 15516822"
  - "Ignas Jasinskis - 15381536"
  - "Radin Azadmanesh - 15486397"
  - "Denis Goncharov - 15654176"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)    # no ## as start of output

rm(list = ls());  cat("\014")          # start with a clean sheet
```


\newpage


# Part 1: Analyzing the Mincerian wage model

## Exercise 1.1

We start by estimating the model.
```{r echo=FALSE}
file = "bankwage_data.csv"
data   = read.csv(file)

y = data$lwage
n = nrow(data)
K = 3
X = cbind(rep(1,n), data$educ, data$exp)

bhat = solve(t(X) %*% X) %*% t(X) %*% y

bhat_vec = setNames(as.vector(bhat), c("Intercept", "Education", "Experience"))
round(bhat_vec, 3)
```
As the estimated OLS coefficients are small, we can conclude that one unit 
change in a regressor $x_{k,i}$ translates to $\beta_k \times 100\%$ change in $y_i$.
In other words, each additional unit of education increases wages by 11.8%,
whereas each additional unit of experience increases wages only by 2.2%.

\newpage

## Exercise 1.2

The correlation coefficient can be measured as
$$r_{\hat{\beta}_2,\hat{\beta}_3} = \frac{Cov(\hat{\beta}_2,\hat{\beta}_3)} {\sqrt{Var(\hat{\beta}_2)Var(\hat{\beta}_3)}}.$$
Computing the expression in R yields the following result:
```{r echo=FALSE}
yhat = X %*% bhat
e = y - yhat
ssr = sum(e*e)
sample_var = ssr/(n-K)
var_cov = sample_var * solve(t(X) %*% X)

cov_b2b3 = var_cov[3,2]
var_b2 = var_cov[2,2]
var_b3 = var_cov[3,3]

corr_b2b3 = cov_b2b3 / sqrt(var_b2 * var_b3)

cat(round(corr_b2b3, 4))
```
Thus, there is small negative correlation between
$Var(\hat{\beta}_2), Var(\hat{\beta}_3)$.


\newpage

## Exercise 1.3
### Part a)

We test non-linearity using the regression specification error test "RESET" of Ramsey. The hypotheses are the following: $H_0:\gamma = 0$ vs $H_a: \gamma \ne 0$ for test equation which is $y_i = x'_i\beta + \gamma \hat{y}_i^2 + \epsilon_i$. The regular estimates rely on the classical OLS variance formula, $\sigma^2(X'X)^{-1}$. The robust estimates, however, use the Eicker-Huber-White 'sandwich' form, $(X'X)^{-1}(X'\Omega X)(X'X)^{-1}$, providing a heteroskedasticity-consistent covariance matrix. After computing the t-test, under $H_0$, for both standard error variations, we get the following results: 
```{r echo = FALSE}
Xa = cbind(X, yhat^2)
Ka = ncol(Xa)

bhat_a = solve(t(Xa) %*% Xa) %*% t(Xa) %*% y
ea     = y - Xa %*% bhat_a
gamma_hat = bhat_a[Ka]

ssr_a      = sum(ea * ea)
sigma2_reg = ssr_a / (n - Ka)
se_reg     = sqrt((sigma2_reg * solve(t(Xa) %*% Xa))[Ka, Ka])
t_stat_reg = gamma_hat / se_reg
p_val_reg  = 2 * (1 - pt(abs(t_stat_reg), n - Ka))

XXinv_a    = solve(t(Xa) %*% Xa)
Sigma_hat     = t(Xa) %*% (diag(as.vector(ea^2))) %*% Xa
vcov_rob_a = XXinv_a %*% Sigma_hat %*% XXinv_a * (n / (n - Ka))
se_rob     = sqrt(vcov_rob_a[Ka, Ka])
t_stat_rob = gamma_hat / se_rob
p_val_rob  = 2 * (1 - pt(abs(t_stat_rob), n - Ka))

df = n - Ka

cat("t-test critical values at different significance levels:")
cat("at 1%  | critical value:", round(qt(1-0.01/2, df), 4), "\n")
cat("at 5%  | critical value:", round(qt(1-0.05/2, df), 4), "\n")
cat("at 10% | critical value:", round(qt(1-0.10/2, df), 4), "\n")

cat("Regular test:", "t-test value:", round(t_stat_reg, 4), " | p-value:", round(p_val_reg, 5), "\n")
cat("Robust test:", "t-test value:", round(t_stat_rob, 4), " | p-value:", round(p_val_rob, 5), "\n")
```
These results conclude that there is no significant evidence to reject the null hypothesis for both calculations - with regular standard errors and with robust standard errors - as t-test values are not in the rejection region and p-values are greater than 1%, 5% or 10% significances.

### Part b)

We want to test for the hypothesis $H_0: \gamma_1 = \gamma_2 = \gamma_3 = 0$ vs $H_a:$ at least one $\gamma_j \ne 0,j=1,2,3$ for the test equation $y_i = x'_i\beta + \gamma_j\Sigma_{j=1}^3 \hat{y}_i^{1+j} + \epsilon_i$. After computing Wald test statistics (df = 3), using regular and robust to heteroscedasicity SE's, we get the following results for 1%, 5% and 10% significance levels:
```{r echo = FALSE}
Xb = cbind(X, yhat^2, yhat^3, yhat^4)
Kb = ncol(Xb)
q  = 3

bhat_b = solve(t(Xb) %*% Xb) %*% t(Xb) %*% y
eb     = y - Xb %*% bhat_b
g_vec  = bhat_b[(Kb - q + 1):Kb] 

ssr_b      = sum(eb * eb)
sigma2_b   = ssr_b / (n - Kb)
vcov_b_reg = sigma2_b * solve(t(Xb) %*% Xb)
V_qq_reg   = vcov_b_reg[(Kb - q + 1):Kb, (Kb - q + 1):Kb]
W_reg      = as.numeric(t(g_vec) %*% solve(V_qq_reg) %*% g_vec)
p_val_reg  = 1 - pchisq(W_reg, df = q)

XXinv_b    = solve(t(Xb) %*% Xb)
Sigma_hat_b     = t(Xb) %*% (diag(as.vector(eb^2))) %*% Xb
vcov_b_rob = XXinv_b %*% Sigma_hat_b %*% XXinv_b * (n / (n - Kb))
V_qq_rob   = vcov_b_rob[(Kb - q + 1):Kb, (Kb - q + 1):Kb]
W_rob      = as.numeric(t(g_vec) %*% solve(V_qq_rob) %*% g_vec)
p_val_rob  = 1 - pchisq(W_rob, df = q)

cat("Wald-test critical values at different significance levels (df=3):\n")
cat("at 1%  | critical value:", round(qchisq(1-0.01, q), 4), "\n")
cat("at 5%  | critical value:", round(qchisq(1-0.05, q), 4), "\n")
cat("at 10% | critical value:", round(qchisq(1-0.10, q), 4), "\n\n")

cat("Regular test: Wald value:", round(W_reg, 4), " | p-value:", round(p_val_reg, 5), "\n")
cat("Robust test:  Wald value:", round(W_rob, 4), " | p-value:", round(p_val_rob, 5), "\n")

```
Again, we conclude that there is not enough evidence to reject the null hypothesis as p-values are above the significance levels and Wald-statistics for both - regular and robust - SE's are not in the rejection region.


\newpage

## Exercise 1.4

We want to test for heteroscedasticity, where we will use Breusch-Pagan test and use the regressors of the original model as z-variables. After following the three steps, the results we get are:
```{r echo=FALSE}
e2 = e^2

bhat_aux = solve(t(X) %*% X) %*% t(X) %*% e2
e2_hat   = X %*% bhat_aux

rss_aux = sum((e2 - e2_hat)^2)
tss_aux = sum((e2 - mean(e2))^2)
r2_aux  = 1 - (rss_aux / tss_aux)

LM_stat = n * r2_aux

p_minus_1 = ncol(X) - 1
p_val_bp  = 1 - pchisq(LM_stat, df = p_minus_1)

cat("Breusch-Pagan Test (LM Statistic):\n")
cat("at 1%  | critical value:", round(qchisq(1-0.01, p_minus_1), 4), "\n")
cat("at 5%  | critical value:", round(qchisq(1-0.05, p_minus_1), 4), "\n")
cat("at 10% | critical value:", round(qchisq(1-0.10, p_minus_1), 4), "\n\n")

cat("Breusch-Pagan Test Results:\n")
cat("LM statistic: ", round(LM_stat, 4), " | p-value:", round(p_val_bp, 5), "\n")

```

Since this p-value (0.00109 or 0.109%) is strictly less than the significance levels of 1%, 5%, and 10%, we reject the null hypothesis $H_0$ of homoskedasticity. By rejecting the null hypothesis, we can conclude that the variance of the error terms is not constant across the observations.

While our current test, using the original regressors educ and exp, confirms the presence of heteroskedasticity, the conclusion is technically sensitive to the choice of z-variables. If the true heteroskedasticity is non-linear or driven by excluded factors, a test using only linear z-variables might lack the power to detect it. However, given our very low p-value, it is highly likely that heteroskedasticity is a robust feature of this dataset regardless of the specific z-variables chosen.

\newpage

# Part 2: Simulating the RESET test and Monte Carlo
                                        
```{r echo=FALSE }
file = "bankwage_data.csv"
cat("We are using Dataset %s", file)


data  = read.csv("bankwage_data.csv")
lwage = data$lwage
educ  = data$educ
exp   = data$exp
n     = length(lwage)

```



```{r echo=FALSE}
# Group 85 (odd) -> Dataset A -> n = 500
# M = 85 %% 5 = 0 -> beta4 = 0.004, x4 = exp^2

educ = data$educ
exp  = data$exp
n    = length(educ)                              # 500

X4    = exp^2   
b4    = 0.004
Xdgp  = cbind(rep(1, n), educ, exp, X4)         # regressors of the true model
beta0 = c(1, 0.1, 0.025, b4)                    # parameters of the true model
sigm0 = 1

cat(sprintf("Part 2 setup: n = %d, beta4 = %.4f, x4 = exp^2\n", n, b4))
```

We generate data from $y_i = 1 + 0.1\, educ_i + 0.025\, exp_i + 0.004\, exp_i^2 + \varepsilon_i$ with $\varepsilon_i \sim N(0,1)$ and $n = `r n`$.

## Exercise 2.1: One sample

```{r echo=FALSE}
set.seed(42) 
eps  = rnorm(n, 0, sigm0)                       # generate disturbances N(0, sigma^2)
ysim = Xdgp %*% beta0 + eps                     # together with X, your dataset is complete

# Estimate the misspecified model: y = b1 + b2*educ + b3*exp + error, considering only linear features

X = cbind(rep(1, n), educ, exp)
k = ncol(X)

XtX       = t(X) %*% X
XXinv     = solve(XtX)
diagXXinv = diag(XXinv)
XXinvX    = XXinv %*% t(X)                       # part of estimator that remains the same

bhat = XXinvX %*% ysim
yhat = X %*% bhat
e    = ysim - X %*% bhat
s2   = sum(e * e) / (n - k)
seb  = sqrt(s2 * diagXXinv)

cat("Estimated coefficients (misspecified model):\n")
results = cbind(bhat, seb)
colnames(results) = c("b_hat", "se(b)")
rownames(results) = c("constant  ", "educ", "exp")
print(results, digits = 5)
```

We analyze the scatter plot of rsiduals against fitted values, distribution of residuals and plot of education agains Y from the DGP. 

```{r echo=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
plot(yhat, e, pch = 20, cex = 0.4, col = "steelblue",
     xlab = "Fitted values", ylab = "Residuals", main = "Residuals vs Fitted")
abline(h = 0, col = 2, lty = 2)


hist(as.numeric(e), breaks = 30, col = "steelblue", main = "Residual Distribution",
     xlab = "Residuals")

#We plot the experience variable against log income to see if there is a non linear pattern there. We fit a second degree polynomial on this scatter plot to see general trend of the variables
plot(exp, ysim, pch = 20, cex = 0.4, col = "steelblue",  
     xlab = "Experience", ylab = "log(Wage)", main = "Experience vs Log(Wage)")
exp_seq  = seq(min(exp), max(exp), length.out = 200)
fit_quad = lm(ysim ~ exp + I(exp^2))
yhat_quad = predict(fit_quad, newdata = data.frame(exp = exp_seq))
lines(exp_seq, yhat_quad, col = "red", lwd = 2)

```
In the figure on the left, we analyze the scatter plot of residuals for different fitted values. We observe approximate randomness of residuals around 0 and no particular patterns emerge, therefore we can not conclude non-linearity of true model.

In the second figure, we see that residuals follow an approximate normal distribution. 

In the third figure, we plot the experience variable against log(wage) to see if there is a non linear pattern there. We fit a second degree polynomial on this scatter plot to see the general trend. Since the fitted line has no visible curve, we can not conclude existence of non-linear experience terms in the model from visual analysis. 
 

### RESET test with $p = 3$

We test the model using RESET test with three additional terms with the model as follows:

$y_i = x'_i\beta + \gamma_j\Sigma_{j=1}^3 \hat{y}_i^{1+j} + \epsilon_i$

With $H_0: \gamma_1 = \gamma_2 = \gamma_3 = 0$ vs $H_a:$ at least one $\gamma_j \ne 0,j=1,2,3$ 

```{r echo=FALSE}

set.seed(0)
# Unrestricted model: add yhat^2, yhat^3, yhat^4
eps  = rnorm(n, 0, sigm0)                    # generate disturbances N(0, sigma^2)
ysim = Xdgp %*% beta0 + eps                  # data from the TRUE model
bhat   = XXinvX %*% ysim   
yhat   = X %*% bhat             # OLS on misspecified model
e      = ysim - yhat


ssrR   = sum(e * e)
Xu     = cbind(X, yhat^2, yhat^3, yhat^4)
ku     = ncol(Xu)                                # 6 regressors in unrestricted model
bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
eu     = ysim - Xu %*% bhat_u
ssrR   = sum(e * e)                              # RSS restricted (original model)
ssrU   = sum(eu * eu)                            # RSS unrestricted (augmented model)
q      = 3                                       # number of restrictions

Fstat  = ((ssrR - ssrU) / q) / (ssrU / (n - ku))

# Robust Wald
XXinvU = solve(t(Xu) %*% Xu)
ku = ncol(Xu)
Sigma_hat = matrix(0, ku, ku)
for (i in 1:nrow(Xu)) {
  xi = Xu[i, ]
  Sigma_hat = Sigma_hat +  eu[i]^2* xi %*% t(xi) 
}
Var_rob   = XXinvU %*% Sigma_hat %*% XXinvU
R      = cbind(matrix(0, q, 3), diag(q))    # We define a constraints such that last three regressors are zero. 
                                              # r = (0,0,0)', so ommitted
Rb_hat     = R %*% bhat_u
Wrob   = (t(Rb_hat) %*% solve(R %*% Var_rob %*% t(R)) %*% Rb_hat)

cat("RESET test (p=3) on simulated data:\n")
cat(sprintf("  F-stat (regular) = %.4f   p-value = %.4f\n", 
            Fstat, 1 - pf(Fstat, q, n - ku)))
cat(sprintf("  Wald (robust)  = %.4f   p-value = %.4f\n", 
            Wrob, pchisq(Wrob, df = 3, lower.tail = FALSE)))
```

We see that using both the simple F-test and robust wald test, the P-value> 0.05, meaning that we can not reject the null hypothesis: 'The model has no omitted variables'. This is likely due to the small coefficient of the exp^2 term in the data generating process. 

\newpage

## Exercise 2.2: 10,000 simulations

In this exercise, we run the same test as before but for 10000 simulations, reporting the rejection rates of the null hypothesis for different significance levels. 

```{r echo=FALSE}
nsim = 10000

F_cv_1  = qf(0.99, 3, n - 6)                      # critical values F(3, n-6)
F_cv_5  = qf(0.95, 3, n - 6)
F_cv_10 = qf(0.90, 3, n - 6)
W_cv_1  = qchisq(0.99, 3)                      # critical values Chisquare(3)
W_cv_5  = qchisq(0.95, 3)
W_cv_10 = qchisq(0.90, 3)
q=3


nrej1 = 0; nrej5 = 0; nrej10 = 0    #Normal rejects
Rnrej1 = 0; Rnrej5 = 0; Rnrej10 = 0  # Robust rejects

set.seed(42)
for (i in 1:nsim) {
  eps  = rnorm(n, 0, sigm0)                    # generate disturbances N(0, sigma^2)
  ysim = Xdgp %*% beta0 + eps                  # data from the TRUE model
  bhat   = XXinvX %*% ysim   
  yhat   = X %*% bhat             # OLS on misspecified model
  e      = ysim - yhat
  ssrR   = sum(e * e)
  
  
  
  Xu     = cbind(X, yhat^2, yhat^3, yhat^4)    # RESET augmentation
  bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
  eu     = ysim - Xu %*% bhat_u
  ssrU   = sum(eu * eu)
  
  
  #Simple F test
  Fstat  = ((ssrR - ssrU) / 3) / (ssrU / (n - 6))
  
  if (Fstat > F_cv_1)  nrej1  = nrej1  + 1
  if (Fstat > F_cv_5)  nrej5  = nrej5  + 1
  if (Fstat > F_cv_10) nrej10 = nrej10 + 1
   
   #Robust test
  XXinvU = solve(t(Xu) %*% Xu)
  ku = ncol(Xu)
  Sigma_hat = matrix(0, ku, ku)
  for (j in 1:nrow(Xu)) {
    xi = Xu[j, ]
    Sigma_hat = Sigma_hat +  eu[j]^2* xi %*% t(xi) 
  }
  Var_rob   = XXinvU %*% Sigma_hat %*% XXinvU
  R      = cbind(matrix(0, q, 3), diag(q))    # We define a constraints such that last three regressors are zero. 
                                                # r = (0,0,0)', so omitted
  Rb_hat     = R %*% bhat_u
  Wrob   = (t(Rb_hat) %*% solve(R %*% Var_rob %*% t(R)) %*% Rb_hat)
  
  if (Wrob > W_cv_1)  Rnrej1  = Rnrej1  + 1
  if (Wrob > W_cv_5)  Rnrej5  = Rnrej5  + 1
  if (Wrob > W_cv_10) Rnrej10 = Rnrej10 + 1
  
}

cat("Rejection rates of simple test over 10000 simulations:\n")
cat(sprintf("  1%%:  %.2f%%\n", nrej1  / nsim * 100))
cat(sprintf("  5%%:  %.2f%%\n", nrej5  / nsim * 100))
cat(sprintf("  10%%: %.2f%%\n", nrej10 / nsim * 100))

cat("Rejection rates of robust test over 10000 simulations:\n")
cat(sprintf("  1%%:  %.2f%%\n", Rnrej1  / nsim * 100))
cat(sprintf("  5%%:  %.2f%%\n",Rnrej5  / nsim * 100))
cat(sprintf("  10%%: %.2f%%\n", Rnrej10 / nsim * 100))
```

The null hypothesis is rejected more often than not, with at least 64.71% times with the robust test with significance level 0.01. At Significance 0.1, the hypothesis is rejected 88.84% of times with robust test and 87.58% with normal F-test. 

\newpage

## Exercise 2.3: Power curve

We vary $\beta_4$ to see how the rejection rate changes. Later, we test for non-linearity using the RESET F-test and the robust Wald test and plot the rejection probability as a function of $\beta_4$. We analyze the rejection rates for $\beta_4$ in range (0,0.01). 


```{r echo=FALSE, fig.height=5, fig.width=7}
b4_range=seq(0, 0.01, by = 0.002)
nsim2= 10000
q= 3                                     # number of restrictions (yhat^2, yhat^3, yhat^4)

rej_F= rep(0, length(b4_range))              # rejection counts for F
rej_W= rep(0, length(b4_range))              # rejection counts for robust Wald

F_cv_5= qf(0.95, q, n - 6)                   # critical value f 5%
W_cv_5= qchisq(0.95, q)                      # critical value chi 5%

R= cbind(matrix(0, q, 3), diag(q))       # constraint: coefficients = 0

set.seed(42)
for (j in 1:length(b4_range)) {
   beta0_j= c(1, 0.1, 0.025, b4_range[j])
   countF = 0
   countW = 0

   for (i in 1:nsim2) {
      eps = rnorm(n, 0, sigm0)
      ysim = Xdgp %*% beta0_j + eps

      # OLS on misspecified model
      bhat= XXinvX %*% ysim
      yhat= X %*% bhat
      e = ysim - yhat
      ssrR = sum(e * e)

      # RESET augmentation
      Xu  = cbind(X, yhat^2, yhat^3, yhat^4)
      bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
      eu= ysim - Xu %*% bhat_u
      ssrU= sum(eu * eu)

      # Regular F-test
      Fstat  = ((ssrR - ssrU) / q) / (ssrU / (n - 6))
      if (Fstat > F_cv_5) countF = countF + 1

      # Robust Wald test
      XXinvU= solve(t(Xu) %*% Xu)
      ku= ncol(Xu)
      Sigma_hat = matrix(0, ku, ku)
      for (l in 1:n) {
         xi = Xu[l, ]
         Sigma_hat = Sigma_hat + eu[l]^2 * xi %*% t(xi)
      }
      Var_rob = XXinvU %*% Sigma_hat %*% XXinvU
      Rb_hat  = R %*% bhat_u
      Wrob  = t(Rb_hat) %*% solve(R %*% Var_rob %*% t(R)) %*% Rb_hat
      if (Wrob > W_cv_5) countW = countW + 1
   }
   rej_F[j] = countF / nsim2
   rej_W[j] = countW / nsim2
}

# Plot both power curves
plot(b4_range, rej_F, type = "b", pch = 20, col = "blueviolet",
     xlab = expression(beta[4]), ylab = "Rejection probability",
     main = "RESET Power Curve (5% level)", ylim = c(0, max(rej_F, rej_W) * 1.1))
lines(b4_range, rej_W, type = "b", pch = 17, col = "orchid3")
abline(h = 0.05, col = 2, lty = 2)
legend("topleft", legend = c("F-test", "Robust Wald"),
       col = c("blueviolet", "orchid3"), pch = c(20, 17), lty = 1)
```

```{r echo=FALSE}
cat("Power curve values (5% level):\n\n")
cat(sprintf("  %-10s  %10s  %12s\n", "beta4", "F-test", "Robust Wald"))
for (j in 1:length(b4_range)) {
   cat(sprintf("  %-10.3f  %9.2f%%  %11.2f%%\n",
       b4_range[j], rej_F[j] * 100, rej_W[j] * 100))
}
```

At $\beta_4 = 0$ the simulated rejection probability should be near 5%. As $\beta_4$ increases, omitting $exp^2$ causes more difference between the true model and the omitted one. The RESET test in that case states about non-linearity more often, so the rejection rises. Power curve shows us how the probability of "correctly" rejecting $H_0$ increases with the size of the omitted variable effect.

\newpage

## Exercise 2.4: Heteroscedastic errors

We introduce heteroscedastic errors: $\varepsilon_i \sim N(0,\; \exp(\alpha \cdot educ_i))$ with $\alpha = 0.1$.

```{r echo=FALSE}
alpha_het = 0.1
nsim3     = 10000

cv_F_1  = qf(0.99, 3, n - 6)
cv_F_5  = qf(0.95, 3, n - 6)
cv_F_10 = qf(0.90, 3, n - 6)
cv_bp_1  = qf(0.99, 4, n - 5)
cv_bp_5  = qf(0.95, 4, n - 5)
cv_bp_10 = qf(0.90, 4, n - 5)

# Counters for RESET
nrej_reset_1 = 0; nrej_reset_5 = 0; nrej_reset_10 = 0
# Counters for BP
nrej_bp_1 = 0; nrej_bp_5 = 0; nrej_bp_10 = 0

set.seed(999)
for (i in 1:nsim3) {
   # Heteroscedastic errors: Var(eps_i) = exp(alpha * educ_i)
   sd_i = exp(alpha_het * educ / 2)     # sd so that Var = exp(alpha * educ)
   eps  = rnorm(n, 0, sd_i)
   ysim = Xdgp %*% beta0 + eps

   bhat   = XXinvX %*% ysim                     # OLS on misspecified model
   e      = ysim - X %*% bhat
   ssrR   = sum(e * e)

   # RESET test
   yhat   = X %*% bhat
   Xu     = cbind(X, yhat^2, yhat^3, yhat^4)
   bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
   eu     = ysim - Xu %*% bhat_u
   ssrU   = sum(eu * eu)
   Fstat  = ((ssrR - ssrU) / 3) / (ssrU / (n - 6))

   if (Fstat > cv_F_1)  nrej_reset_1  = nrej_reset_1  + 1
   if (Fstat > cv_F_5)  nrej_reset_5  = nrej_reset_5  + 1
   if (Fstat > cv_F_10) nrej_reset_10 = nrej_reset_10 + 1

  # Testing for heteroscedasticity 
  # We create regression of different combinations of residuals (Using White's statistic) and check if X explains       residuals
  
  X_bp = cbind(rep(1, n),educ,exp,educ^2,exp^2)
  
  gamma_hat = solve(t(X_bp) %*% X_bp, t(X_bp) %*% e^2) # We use e, which is residuals from restricted model
  e_bp = e^2 - X_bp %*% gamma_hat
  SSR_bp = sum(e_bp*e_bp)
  SST_bp = sum((e^2-mean(e^2))^2)
  R2_bp= (1-(SSR_bp/SST_bp))
  Fstat_bp  = (R2_bp / 4) / ((1-R2_bp) / (n - 5)) #Fstat where we have p-1=4 features (including the squares)
  
  
  if (Fstat_bp > cv_bp_1)  nrej_bp_1  = nrej_bp_1  + 1
  if (Fstat_bp > cv_bp_5)  nrej_bp_5  = nrej_bp_5  + 1
  if (Fstat_bp > cv_bp_10) nrej_bp_10 = nrej_bp_10 + 1

   

   
}

cat("With heteroscedastic errors (alpha = 0.1):\n\n")
cat("RESET test rejection rates:\n")
cat(sprintf("  1%%: %.2f%%   5%%: %.2f%%   10%%: %.2f%%\n",
    nrej_reset_1 / nsim3 * 100, nrej_reset_5 / nsim3 * 100,
    nrej_reset_10 / nsim3 * 100))
cat("\nBreusch-Pagan test rejection rates:\n")
cat(sprintf("  1%%: %.2f%%   5%%: %.2f%%   10%%: %.2f%%\n",
    nrej_bp_1 / nsim3 * 100, nrej_bp_5 / nsim3 * 100,
    nrej_bp_10 / nsim3 * 100))
```

With heteroscedastic errors, the Breusch--Pagan test should reject $H_0$ much more frequently than the nominal level, confirming its power against heteroscedasticity. The RESET test may also show elevated rejection rates since heteroscedasticity inflates the F-statistic when regular S.E. are used.