---
title:  "Assignment 2"
author: ""
date:   "14/03/2026"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)    # no ## as start of output

rm(list = ls());  cat("\014")          # start with a clean sheet
```

## Part 1

You might develop part 1 and part 2 separately, but in the end it should be combined and kitted into one pdf.
Be aware that this will only succeed if ALL chunks are error-free.

***

The chunk below can be used for reading the data file and setting up the variables needed for estimating the model. You might even include some basic statistics and (scatter)plots.

```{r echo=FALSE}

for (i in 1:5) cat(sprintf(".\n"))
   



```


RMarkdown sections can be used, either to comment on the {r} chunk above, 

---

or to introduce the next {r} chunk.
 
                                        
```{r echo=TRUE}
#
#
# For parts of the code that you want in the knitted file.
#
#


```
\newpage

## Part 2

In the chunk below, a matrix is constructed containing a column with ones and two regressors taken from the given data files.
                                        
```{r echo=FALSE }
path = "C:/Users/...."

file = "..."
sprintf("We are using Dataset %s", file)
data   = read.csv(file)

#  extract the variables educ and exp

n    = length(educ)
Xdgp = cbind( rep(1,n), educ, exp, X4)        # regressors of the true model
beta0 = c(1, 0.1, 0.025, b4)                  # parameters of the true model
sigm0 = 1

```
\newpage 
## Exercise 2.1
### One sample of $n = 500$

```{r echo=TRUE }
set.seed(100) 
eps      = rnorm(n, 0, sigm0)                       # generate disturbances N(0, sigma^2)
ysim     = Xdgp %*% beta0 + eps                     # together with X, your data set is complete

#
#  Code for exercise 2.1
#

X = cbind( rep(1,n), , , ,  )          # setup the model you want to estimate
k = ncol(X)

XtX   = t(X) %*% X                     # matrix with basic statistics (n, sum(x), sum(x^2))
XXinv = solve(XtX)                     # matrix used in Variance of estimator b
diagXXinv = diag(XXinv)                # the diagonal elements (for the variance of b)
XXinvX    = XXinv %*% t(X)             # part of the estimator betahat that remains the same





```


```{r echo=TRUE }
#
#  More code for exercise 2.1 (RESET test)
#





```



```{r echo=TRUE }
#
#  Code for exercise 2.2
#



for (i in 1:nsim) {





#
#  More code for exercise 2.2
#






}
```


For the next questions, copy/paste the code from the previous exercise and make the required adjustments.
