---
title: "Untitled"
output: html_document
date: "2026-03-18"
---

```{r echo=FALSE}
data  = read.csv("bankwage_data.csv")
lwage = data$lwage
educ  = data$educ
exp   = data$exp
n     = length(lwage)
```

\newpage

# Part 2: Simulate the RESET test

```{r echo=FALSE}
# ---- DGP Setup ----
# Group 85 (odd) -> Dataset A -> n = 500
# M = 85 %% 5 = 0 -> beta4 = 0.004, x4 = exp^2

educ = data$educ
exp  = data$exp
n    = length(educ)                              # 500

X4    = exp^2   
b4    = 0.004
Xdgp  = cbind(rep(1, n), educ, exp, X4)         # regressors of the true model
beta0 = c(1, 0.1, 0.025, b4)                    # parameters of the true model
sigm0 = 1

cat(sprintf("Part 2 setup: n = %d, beta4 = %.4f, x4 = exp^2\n", n, b4))
```

We generate data from $y_i = 1 + 0.1\, educ_i + 0.025\, exp_i + 0.004\, exp_i^2 + \varepsilon_i$ with $\varepsilon_i \sim N(0,1)$ and $n = `r n`$.

## Exercise 2.1: One sample

```{r echo=TRUE}
set.seed(42) 
eps  = rnorm(n, 0, sigm0)                       # generate disturbances N(0, sigma^2)
ysim = Xdgp %*% beta0 + eps                     # together with X, your dataset is complete

# Estimate the misspecified model: y = b1 + b2*educ + b3*exp + error, considering only linear features

X = cbind(rep(1, n), educ, exp)
k = ncol(X)

XtX       = t(X) %*% X
XXinv     = solve(XtX)
diagXXinv = diag(XXinv)
XXinvX    = XXinv %*% t(X)                       # part of estimator that remains the same

bhat = XXinvX %*% ysim
yhat = X %*% bhat
e    = ysim - X %*% bhat
s2   = sum(e * e) / (n - k)
seb  = sqrt(s2 * diagXXinv)

cat("Estimated coefficients (misspecified model):\n")
results = cbind(bhat, seb)
colnames(results) = c("b_hat", "se(b)")
rownames(results) = c("constant  ", "educ", "exp")
print(results, digits = 5)
```

 

### RESET test with $p = 3$

```{r echo=TRUE}
# Unrestricted model: add yhat^2, yhat^3, yhat^4
Xu     = cbind(X, yhat^2, yhat^3, yhat^4)
ku     = ncol(Xu)                                # 6 regressors in unrestricted model
bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
eu     = ysim - Xu %*% bhat_u
ssrR   = sum(e * e)                              # RSS restricted (original model)
ssrU   = sum(eu * eu)                            # RSS unrestricted (augmented model)
q      = 3                                       # number of restrictions

Fstat  = ((ssrR - ssrU) / q) / (ssrU / (n - ku))

# Robust Wald
XXinvU = solve(t(Xu) %*% Xu)
XXinvU = solve(t(Xu) %*% Xu)
ku = ncol(Xu)
Sigma_hat = matrix(0, ku, ku)
for (i in 1:nrow(Xu)) {
  xi = Xu[i, ]
  Sigma_hat = Sigma_hat +  e[i]^2* xi %*% t(xi) 
}
Var_rob   = XXinvU %*% Sigma_hat %*% XXinvU
R      = cbind(matrix(0, q, 3), diag(q))    # We define a constraints such that last three regressors are zero. 
                                              # r = (0,0,0)', so ommitted
Rb_hat     = R %*% bhat_u
Wrob   = (t(Rb_hat) %*% solve(R %*% Var_rob %*% t(R)) %*% Rb_hat)

cat("RESET test (p=3) on simulated data:\n")
cat(sprintf("  F-stat (regular) = %.4f   p-value = %.4f\n", 
            Fstat, 1 - pf(Fstat, q, n - ku)))
cat(sprintf("  Wald (robust)  = %.4f   p-value = %.4f\n", 
            Wrob, pchisq(Wrob, df = 3, lower.tail = FALSE)))
```

We see that using both the simple F-test and robust wald test, the P-value> 0.05, meaning that we reject the null hypothesis: 'The model has no ommitted variables'
\newpage

## Exercise 2.2: 10,000 simulations

```{r echo=TRUE}
nsim = 10000

F_cv_1  = qf(0.99, 3, n - 6)                      # critical values F(3, n-6)
F_cv_5  = qf(0.95, 3, n - 6)
F_cv_10 = qf(0.90, 3, n - 6)
W_cv_1  = qchisq(0.99, 3)                      # critical values Chisquare(3)
W_cv_5  = qchisq(0.95, 3)
W_cv_10 = qchisq(0.90, 3)


nrej1 = 0; nrej5 = 0; nrej10 = 0    #Normal rejects
Rnrej1 = 0; Rnrej5 = 0; Rnrej10 = 0  # Robust rejects

set.seed(42)
for (i in 1:nsim) {
  eps  = rnorm(n, 0, sigm0)                    # generate disturbances N(0, sigma^2)
  ysim = Xdgp %*% beta0 + eps                  # data from the TRUE model
  bhat   = XXinvX %*% ysim   
  yhat   = X %*% bhat             # OLS on misspecified model
  e      = ysim - yhat
  ssrR   = sum(e * e)
  
  
  Xu     = cbind(X, yhat^2, yhat^3, yhat^4)    # RESET augmentation
  bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
  eu     = ysim - Xu %*% bhat_u
  ssrU   = sum(eu * eu)
  
  
  #Simple F test
  Fstat  = ((ssrR - ssrU) / 3) / (ssrU / (n - 6))
  
  if (Fstat > F_cv_1)  nrej1  = nrej1  + 1
  if (Fstat > F_cv_5)  nrej5  = nrej5  + 1
  if (Fstat > F_cv_10) nrej10 = nrej10 + 1
   
   #Robust test
  XXinvU = solve(t(Xu) %*% Xu)
  XXinvU = solve(t(Xu) %*% Xu)
  ku = ncol(Xu)
  Sigma_hat = matrix(0, ku, ku)
  for (i in 1:nrow(Xu)) {
    xi = Xu[i, ]
    Sigma_hat = Sigma_hat +  eu[i]^2* xi %*% t(xi) 
  }
  Var_rob   = XXinvU %*% Sigma_hat %*% XXinvU
  R      = cbind(matrix(0, q, 3), diag(q))    # We define a constraints such that last three regressors are zero. 
                                                # r = (0,0,0)', so omitted
  Rb_hat     = R %*% bhat_u
  Wrob   = (t(Rb_hat) %*% solve(R %*% Var_rob %*% t(R)) %*% Rb_hat)
  
  if (Wrob > W_cv_1)  Rnrej1  = Rnrej1  + 1
  if (Wrob > W_cv_5)  Rnrej5  = Rnrej5  + 1
  if (Wrob > W_cv_10) Rnrej10 = Rnrej10 + 1
  
}

cat("Rejection rates over 10000 simulations:\n")
cat(sprintf("  1%%:  %.2f%%\n", nrej1  / nsim * 100))
cat(sprintf("  5%%:  %.2f%%\n", nrej5  / nsim * 100))
cat(sprintf("  10%%: %.2f%%\n", nrej10 / nsim * 100))

cat("Rejection rates over 10000 simulations:\n")
cat(sprintf("  1%%:  %.2f%%\n", Rnrej1  / nsim * 100))
cat(sprintf("  5%%:  %.2f%%\n",Rnrej5  / nsim * 100))
cat(sprintf("  10%%: %.2f%%\n", Rnrej10 / nsim * 100))
```

The null hypothesis is rejected very often, with at least 54.33% times with the robust test with significance level 0.01. At Significance 0.1, the hypothesis is rejected 85.95% of times with robust test and 87.58% with normal F-test. 

\newpage

## Exercise 2.3: Power curve

We vary $\beta_4$ to see how the rejection rate changes. Later, we test for non-linearity using the RESET F-test and the robust Wald test and plot the rejection probability as a function of $\beta_4$.


```{r echo=TRUE, fig.height=5, fig.width=7}
b4_range=seq(0, 0.01, by = 0.002)
nsim2= 10000
q= 3                                     # number of restrictions (yhat^2, yhat^3, yhat^4)

rej_F= rep(0, length(b4_range))              # rejection counts for F
rej_W= rep(0, length(b4_range))              # rejection counts for robust Wald

F_cv_5= qf(0.95, q, n - 6)                   # critical value f 5%
W_cv_5= qchisq(0.95, q)                      # critical value chi 5%

R= cbind(matrix(0, q, 3), diag(q))       # constraint: coefficients = 0

set.seed(42)
for (j in 1:length(b4_range)) {
   beta0_j= c(1, 0.1, 0.025, b4_range[j])
   countF = 0
   countW = 0

   for (i in 1:nsim2) {
      eps = rnorm(n, 0, sigm0)
      ysim = Xdgp %*% beta0_j + eps

      # OLS on misspecified model
      bhat= XXinvX %*% ysim
      yhat= X %*% bhat
      e = ysim - yhat
      ssrR = sum(e * e)

      # RESET augmentation
      Xu  = cbind(X, yhat^2, yhat^3, yhat^4)
      bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
      eu= ysim - Xu %*% bhat_u
      ssrU= sum(eu * eu)

      # Regular F-test
      Fstat  = ((ssrR - ssrU) / q) / (ssrU / (n - 6))
      if (Fstat > F_cv_5) countF = countF + 1

      # Robust Wald test
      XXinvU= solve(t(Xu) %*% Xu)
      ku= ncol(Xu)
      Sigma_hat = matrix(0, ku, ku)
      for (l in 1:n) {
         xi = Xu[l, ]
         Sigma_hat = Sigma_hat + eu[l]^2 * xi %*% t(xi)
      }
      Var_rob = XXinvU %*% Sigma_hat %*% XXinvU
      Rb_hat  = R %*% bhat_u
      Wrob  = t(Rb_hat) %*% solve(R %*% Var_rob %*% t(R)) %*% Rb_hat
      if (Wrob > W_cv_5) countW = countW + 1
   }
   rej_F[j] = countF / nsim2
   rej_W[j] = countW / nsim2
}

# Plot both power curves
plot(b4_range, rej_F, type = "b", pch = 20, col = "blueviolet",
     xlab = expression(beta[4]), ylab = "Rejection probability",
     main = "RESET Power Curve (5% level)", ylim = c(0, max(rej_F, rej_W) * 1.1))
lines(b4_range, rej_W, type = "b", pch = 17, col = "orchid3")
abline(h = 0.05, col = 2, lty = 2)
legend("topleft", legend = c("F-test", "Robust Wald"),
       col = c("blueviolet", "orchid3"), pch = c(20, 17), lty = 1)
```

```{r echo=TRUE}
cat("Power curve values (5% level):\n\n")
cat(sprintf("  %-10s  %10s  %12s\n", "beta4", "F-test", "Robust Wald"))
for (j in 1:length(b4_range)) {
   cat(sprintf("  %-10.3f  %9.2f%%  %11.2f%%\n",
       b4_range[j], rej_F[j] * 100, rej_W[j] * 100))
}
```

At $\beta_4 = 0$ the simulated rejection probability should be near 5%. As $\beta_4$ increases, omitting $exp^2$ causes more difference between the true model and the omitted one. The RESET test in that case states about non-linearity more often, so the rejection rises. Power curve shows us how the probability of "correctly" rejecting $H_0$ increases with the size of the omitted variable effect.

\newpage

## Exercise 2.4: Heteroscedastic errors

We introduce heteroscedastic errors: $\varepsilon_i \sim N(0,\; \exp(\alpha \cdot educ_i))$ with $\alpha = 0.1$.

```{r echo=TRUE}
alpha_het = 0.1
nsim3     = 10000

cv_F_1  = qf(0.99, 3, n - 6)
cv_F_5  = qf(0.95, 3, n - 6)
cv_F_10 = qf(0.90, 3, n - 6)
cv_bp_1  = qf(0.99, 4, n - 5)
cv_bp_5  = qf(0.95, 4, n - 5)
cv_bp_10 = qf(0.90, 4, n - 5)

# Counters for RESET
nrej_reset_1 = 0; nrej_reset_5 = 0; nrej_reset_10 = 0
# Counters for BP
nrej_bp_1 = 0; nrej_bp_5 = 0; nrej_bp_10 = 0

set.seed(999)
for (i in 1:nsim3) {
   # Heteroscedastic errors: Var(eps_i) = exp(alpha * educ_i)
   sd_i = exp(alpha_het * educ / 2)     # sd so that Var = exp(alpha * educ)
   eps  = rnorm(n, 0, sd_i)
   ysim = Xdgp %*% beta0 + eps

   bhat   = XXinvX %*% ysim                     # OLS on misspecified model
   e      = ysim - X %*% bhat
   ssrR   = sum(e * e)

   # ---- RESET test ----
   yhat   = X %*% bhat
   Xu     = cbind(X, yhat^2, yhat^3, yhat^4)
   bhat_u = solve(t(Xu) %*% Xu, t(Xu) %*% ysim)
   eu     = ysim - Xu %*% bhat_u
   ssrU   = sum(eu * eu)
   Fstat  = ((ssrR - ssrU) / 3) / (ssrU / (n - 6))

   if (Fstat > cv_F_1)  nrej_reset_1  = nrej_reset_1  + 1
   if (Fstat > cv_F_5)  nrej_reset_5  = nrej_reset_5  + 1
   if (Fstat > cv_F_10) nrej_reset_10 = nrej_reset_10 + 1

  # Testing for heteroscedasticity 
  # We create regression of different combinations of residuals (Using White's statistic) and check if X explains       residuals
  
  X_bp = cbind(rep(1, n),educ,exp,educ^2,exp^2)
  
  gamma_hat = solve(t(X_bp) %*% X_bp, t(X_bp) %*% e^2) # We use e, which is residuals from restricted model
  e_bp = e^2 - X_bp %*% gamma_hat
  SSR_bp = sum(e_bp*e_bp)
  SST_bp = sum((e^2-mean(e^2))^2)
  R2_bp= (1-(SSR_bp/SST_bp))
  Fstat_bp  = (R2_bp / 4) / ((1-R2_bp) / (n - 5)) #Fstat where we have p-1=4 features (including the squares)
  
  
  if (Fstat_bp > cv_bp_1)  nrej_bp_1  = nrej_bp_1  + 1
  if (Fstat_bp > cv_bp_5)  nrej_bp_5  = nrej_bp_5  + 1
  if (Fstat_bp > cv_bp_10) nrej_bp_10 = nrej_bp_10 + 1

   

   
}

cat("With heteroscedastic errors (alpha = 0.1):\n\n")
cat("RESET test rejection rates:\n")
cat(sprintf("  1%%: %.2f%%   5%%: %.2f%%   10%%: %.2f%%\n",
    nrej_reset_1 / nsim3 * 100, nrej_reset_5 / nsim3 * 100,
    nrej_reset_10 / nsim3 * 100))
cat("\nBreusch-Pagan test rejection rates:\n")
cat(sprintf("  1%%: %.2f%%   5%%: %.2f%%   10%%: %.2f%%\n",
    nrej_bp_1 / nsim3 * 100, nrej_bp_5 / nsim3 * 100,
    nrej_bp_10 / nsim3 * 100))
```

With heteroscedastic errors, the Breusch--Pagan test should reject $H_0$ much more frequently than the nominal level, confirming its power against heteroscedasticity. The RESET test may also show elevated rejection rates since heteroscedasticity inflates the F-statistic when regular S.E. are used.
