---
title:  "Assignment 2"
authors: "Dovydas Venclovas, Ignas Jasinskis, Radin Azadmanesh, Denis Goncharov"
date:   "20/03/2026"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)    # no ## as start of output

rm(list = ls());  cat("\014")          # start with a clean sheet
```

## Part 1

## Exercise 1.1

We start by estimating the model.
```{r echo=FALSE}
file = "bankwage_data.csv"
data   = read.csv(file)

y = data$lwage
n = nrow(data)
K = 3
X = cbind(rep(1,n), data$educ, data$exp)

bhat = solve(t(X) %*% X) %*% t(X) %*% y

bhat_vec = setNames(as.vector(bhat), c("Intercept", "Education", "Experience"))
round(bhat_vec, 3)
```
As the estimated OLS coefficients are small, we can conclude that one unit 
change in a regressor $x_{k,i}$ translates to $\beta_k \times 100\%$ change in $y_i$.
In other words, each additional unit of education increases wages by 11.8%,
whereas each additional unit of experience increases wages only by 2.2%.

## Exercise 1.2

The correlation coefficient can be measured as
$$r_{\hat{\beta}_2,\hat{\beta}_3} = \frac{Cov(\hat{\beta}_2,\hat{\beta}_3)} {\sqrt{Var(\hat{\beta}_2)Var(\hat{\beta}_3)}}.$$
Computing the expression in R yields the following result.
```{r echo=FALSE}
e = y - (y - X %*% bhat)
ssr = sum(e*e)
sample_var = ssr/(n-K)
var_cov = sample_var * solve(t(X) %*% X)

cov_b2b3 = var_cov[3,2]
var_b2 = var_cov[2,2]
var_b3 = var_cov[3,3]

corr_b2b3 = cov_b2b3 / sqrt(var_b2 * var_b3)
round(corr_b2b3, 4)
```
\newpage

## Part 2

In the chunk below, a matrix is constructed containing a column with ones and two regressors taken from the given data files.
                                        
```{r echo=FALSE }
# file = "bankwage_data.csv"
# sprintf("We are using Dataset %s", file)
# data   = read.csv(file)

#  extract the variables educ and exp

#n    = length(educ)
#Xdgp = cbind( rep(1,n), educ, exp, X4)        # regressors of the true model
#beta0 = c(1, 0.1, 0.025, b4)                  # parameters of the true model
#sigm0 = 1

```
\newpage 
## Exercise 2.1
### One sample of $n = 500$

```{r echo=TRUE }
#set.seed(100) 
#eps      = rnorm(n, 0, sigm0)                       # generate disturbances N(0, sigma^2)
#ysim     = Xdgp %*% beta0 + eps                     # together with X, your data set is complete

#
#  Code for exercise 2.1
#

#X = cbind( rep(1,n), , , ,  )          # setup the model you want to estimate
#k = ncol(X)

#XtX   = t(X) %*% X                     # matrix with basic statistics (n, sum(x), sum(x^2))
#XXinv = solve(XtX)                     # matrix used in Variance of estimator b
#diagXXinv = diag(XXinv)                # the diagonal elements (for the variance of b)
#XXinvX    = XXinv %*% t(X)             # part of the estimator betahat that remains the same





```


```{r echo=TRUE }
#
#  More code for exercise 2.1 (RESET test)
#





```



```{r echo=TRUE }
#
#  Code for exercise 2.2
#



# for (i in 1:nsim) {





#
#  More code for exercise 2.2
#


# }
```


For the next questions, copy/paste the code from the previous exercise and make the required adjustments.
