---
title:  "Assignment 2"
authors: "Dovydas Venclovas, Ignas Jasinskis, Radin Azadmanesh, Denis Goncharov"
date:   "20/03/2026"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)    # no ## as start of output

rm(list = ls());  cat("\014")          # start with a clean sheet
```

## Part 1

## Exercise 1.1

We start by estimating the model.
```{r echo=FALSE}
file = "bankwage_data.csv"
data   = read.csv(file)

y = data$lwage
n = nrow(data)
K = 3
X = cbind(rep(1,n), data$educ, data$exp)

bhat = solve(t(X) %*% X) %*% t(X) %*% y

bhat_vec = setNames(as.vector(bhat), c("Intercept", "Education", "Experience"))
round(bhat_vec, 3)
```
As the estimated OLS coefficients are small, we can conclude that one unit 
change in a regressor $x_{k,i}$ translates to $\beta_k \times 100\%$ change in $y_i$.
In other words, each additional unit of education increases wages by 11.8%,
whereas each additional unit of experience increases wages only by 2.2%.

## Exercise 1.2

The correlation coefficient can be measured as
$$r_{\hat{\beta}_2,\hat{\beta}_3} = \frac{Cov(\hat{\beta}_2,\hat{\beta}_3)} {\sqrt{Var(\hat{\beta}_2)Var(\hat{\beta}_3)}}.$$
Computing the expression in R yields the following result.
```{r echo=FALSE}
yhat = X %*% bhat
e = y - yhat
ssr = sum(e*e)
sample_var = ssr/(n-K)
var_cov = sample_var * solve(t(X) %*% X)

cov_b2b3 = var_cov[3,2]
var_b2 = var_cov[2,2]
var_b3 = var_cov[3,3]

corr_b2b3 = cov_b2b3 / sqrt(var_b2 * var_b3)
round(corr_b2b3, 4)
```

## Exercise 1.3
## Part a)
```{r echo = FALSE}
Xa = cbind(X, yhat^2)
Ka = ncol(Xa)

bhat_a = solve(t(Xa) %*% Xa) %*% t(Xa) %*% y
ea     = y - Xa %*% bhat_a
gamma_hat = bhat_a[Ka]

ssr_a      = sum(ea * ea)
sigma2_reg = ssr_a / (n - Ka)
se_reg     = sqrt((sigma2_reg * solve(t(Xa) %*% Xa))[Ka, Ka])
t_stat_reg = gamma_hat / se_reg
p_val_reg  = 2 * (1 - pt(abs(t_stat_reg), n - Ka))

XXinv_a    = solve(t(Xa) %*% Xa)
meat_a     = t(Xa) %*% (diag(as.vector(ea^2))) %*% Xa
vcov_rob_a = XXinv_a %*% meat_a %*% XXinv_a * (n / (n - Ka))
se_rob     = sqrt(vcov_rob_a[Ka, Ka])
t_stat_rob = gamma_hat / se_rob
p_val_rob  = 2 * (1 - pt(abs(t_stat_rob), n - Ka))

df = n - Ka

cat("t-test critical values at different significance levels:")
cat("at 1%  | critical value:", round(qt(1-0.01/2, df), 4), "\n")
cat("at 5%  | critical value:", round(qt(1-0.05/2, df), 4), "\n")
cat("at 10% | critical value:", round(qt(1-0.10/2, df), 4), "\n")

cat("Regular test:", "t-test value:", round(t_stat_reg, 4), " | p-value:", round(p_val_reg, 5), "\n")

cat("Robust test:", "t-test value:", round(t_stat_rob, 4), " | p-value:", round(p_val_rob, 5), "\n")

```

## Part b)
```{r echo = FALSE}
Xb = cbind(X, yhat^2, yhat^3, yhat^4)
Kb = ncol(Xb)
q  = 3

bhat_b = solve(t(Xb) %*% Xb) %*% t(Xb) %*% y
eb     = y - Xb %*% bhat_b
g_vec  = bhat_b[(Kb - q + 1):Kb] 

ssr_b      = sum(eb * eb)
sigma2_b   = ssr_b / (n - Kb)
vcov_b_reg = sigma2_b * solve(t(Xb) %*% Xb)
V_qq_reg   = vcov_b_reg[(Kb - q + 1):Kb, (Kb - q + 1):Kb]
W_reg      = as.numeric(t(g_vec) %*% solve(V_qq_reg) %*% g_vec)
p_val_reg  = 1 - pchisq(W_reg, df = q)

XXinv_b    = solve(t(Xb) %*% Xb)
meat_b     = t(Xb) %*% (diag(as.vector(eb^2))) %*% Xb
vcov_b_rob = XXinv_b %*% meat_b %*% XXinv_b * (n / (n - Kb))
V_qq_rob   = vcov_b_rob[(Kb - q + 1):Kb, (Kb - q + 1):Kb]
W_rob      = as.numeric(t(g_vec) %*% solve(V_qq_rob) %*% g_vec)
p_val_rob  = 1 - pchisq(W_rob, df = q)

cat("Wald-test critical values at different significance levels (df=3):\n")
cat("at 1%  | critical value:", round(qchisq(1-0.01, q), 4), "\n")
cat("at 5%  | critical value:", round(qchisq(1-0.05, q), 4), "\n")
cat("at 10% | critical value:", round(qchisq(1-0.10, q), 4), "\n\n")

cat("Regular test: Wald value:", round(W_reg, 4), " | p-value:", round(p_val_reg, 5), "\n")
cat("Robust test:  Wald value:", round(W_rob, 4), " | p-value:", round(p_val_rob, 5), "\n")

```


\newpage

## Part 2

In the chunk below, a matrix is constructed containing a column with ones and two regressors taken from the given data files.
                                        
```{r echo=FALSE }
# file = "bankwage_data.csv"
# sprintf("We are using Dataset %s", file)
# data   = read.csv(file)

#  extract the variables educ and exp

#n    = length(educ)
#Xdgp = cbind( rep(1,n), educ, exp, X4)        # regressors of the true model
#beta0 = c(1, 0.1, 0.025, b4)                  # parameters of the true model
#sigm0 = 1

```
\newpage 
## Exercise 2.1
### One sample of $n = 500$

```{r echo=TRUE }
#set.seed(100) 
#eps      = rnorm(n, 0, sigm0)                       # generate disturbances N(0, sigma^2)
#ysim     = Xdgp %*% beta0 + eps                     # together with X, your data set is complete

#
#  Code for exercise 2.1
#

#X = cbind( rep(1,n), , , ,  )          # setup the model you want to estimate
#k = ncol(X)

#XtX   = t(X) %*% X                     # matrix with basic statistics (n, sum(x), sum(x^2))
#XXinv = solve(XtX)                     # matrix used in Variance of estimator b
#diagXXinv = diag(XXinv)                # the diagonal elements (for the variance of b)
#XXinvX    = XXinv %*% t(X)             # part of the estimator betahat that remains the same





```


```{r echo=TRUE }
#
#  More code for exercise 2.1 (RESET test)
#





```



```{r echo=TRUE }
#
#  Code for exercise 2.2
#



# for (i in 1:nsim) {





#
#  More code for exercise 2.2
#


# }
```


For the next questions, copy/paste the code from the previous exercise and make the required adjustments.
